\textbf{Empirical Bayes shrinkage, and denoising of Poisson and heteroskedastic Gaussian signals (?)}

### 2016-02-15 paper/draft_6.tex
added new files

Methods for smoothing the variance of Gaussian data, or even for smoothing the mean in the presence of heteroskedastic errors, are uncommon, particularly compared with the well-studied homoskedastic case. Previous non-wavelet-based work includes \cite{Fan1998Efficient}, who estimated the variance by smoothing the squared residuals using local polynomial smoothing, and \cite{Brown2007Variance} who employed difference-based kernel estimators. The only wavelet-based method for variance estimation we have found is \cite{Cai2008Adaptive}, which applied a wavelet thresholding approach to first order differences. Unfortunately none of these methods have available software implementations. (WAND'S WORK?)

@@

%In this paper we use \ash~to tackle several signal denoising problems,  including Poisson smoothing,
%and smoothing both the mean and variance of heteroskedastic Gaussian data. 
Shrinkage methods are widely-used in signal denoising applications, because signal denoising can be accurately and conveniently achieved by shrinkage in a transformed (e.g.~wavelet) domain \cite{Donoho1994Ideal}. 
Commonly-used shrinkage methods include both
simple thresholding rules \cite{Coifman1995Translationinvariant, Donoho1994Ideal, donoho95} and EB methods \cite{Johnstone2005Empirical, Clyde2000Flexible}.
Being an EB method, \ash~has much in common with these previous EB methods, but generalizes them in two ways: first, 
\ash~allows more flexibility in the underlying distribution $g$ than the Laplace or spike-and-slab
distributions used in \cite{Johnstone2005Empirical, Clyde2000Flexible}; second \ash~allows for variations in precision in the transformed observations (e.g. wavelet coefficients). The latter property is particularly important for
the Poisson and heteroskedastic Gaussian settings we consider here.

@@

The above outlines the basic strategy, but there are some important
 additional implementational details:
\begin{itemize}
\item Rather than use a single wavelet transform, we use the ``non-decimated" wavelet transform, which treats the observations as coming
from a circle, rather than a line, and averages results over all $T$ possible rotations of the data. Although not always necessary, this is a standard trick to reduce artifacts that can occur near discontinuities in the underlying signal (see eg. Coifman \& Donoho (1995)). 
%COMMENT: WHat about reflection before doing this? ans: all the test fns are periodic
%DO OTHER EB APPROACHES CONSIDER THE NDWT? yes ebayesthresh software discusses this option
\item The non-decimated wavelet transform yields $T$ wavelet coefficients (transformed values of $\bY$) 
at each of $J=\log_2(T)$ resolution levels.
We apply ash separately to the wavelet coefficients at each resolution level, so that
a different distribution $g$ for the $(\tmu_j)$ is estimated for each resolution. This is the usual way that
EB approaches are applied in this context (e.g.~\cite{checkcgandjs}) and indeed is crucial because
the underlying distribution $g$  will vary with resolution (because smoothness of $\bmu$ will vary with resolution).  
\item Although we have presented the wavelet transform as a matrix multiplication, which is an $o(T^2)$ operation, in practice both the wavelet transform and the inverse tranform are implemented using
efficient algorithms \cite{beylkin.92,coifman.donoho.94}, implemented in the {\tt wavethresh} package \cite{}, taking only $O(T\log T)$ operations. 
\end{itemize}
%\item Ash yields a posterior distribution for each $\tmu_t$. A simple approach would
%be to use the posterior mean for $\tmu_t$ and then invert the wavelet transform; however we instead use the delta method?

@@

Although we assumed variances to be known above, the problem of variance estimation is itself non-trivial. 
A common approach is to assume that the variance is constant ($\sigma_t=\sigma$), and that changes in the adjacent means $\mu_t-\mu_{t+1}$ are negligible, and to use an unbiased estimate for $\sigma$ such as
\begin{equation} \label{eq:constvar}
\hat{\sigma}^2 = (1/2T) \sum_{t=1}^{T} (Y_t-Y_{t-1})^2
\end{equation}
(with $Y_0:=Y_T$.
Here we make the more flexible assumption that the variance function is spatially structured, and use wavelet shrinkage to estimate it. Our approach is similar in spirit to \cite{cai.wang}, but we jointly estimate the mean and variance functions whereas they estimate only the variance function.

@@

To extend this to $T=4$, introduce the notation $v_{i:j}$ to denote, for any vector $v$, 
the sum $\sum_{t=i}^{t=j} v_t$.  Then 
\begin{align} 
Y_{1:4} & \sim \Poi(\mu_{1:4}) \label{eqn:y14} \\
Y_{1:2} | Y_{1:4} & \sim \Bin(Y_{1:4}, \mu_{1:2}/\mu_{1:4})  \label{eqn:y12} \\
Y_1 | Y_{1:2} & \sim \Bin(Y_{1:2}, \mu_1/\mu_{1:2}) \label{eqn:y1} \\ 
Y_3 | Y_{3:4} & \sim \Bin(Y_{3:4}, \mu_3/\mu_{3:4}). \label{eqn:y3}
\end{align} 
Together these models are exactly equivalent to $Y_j \sim \Poi(\mu_j)$, and
they decompose the overall distribution $Y_1,\dots,Y_4$ into parts involving aspects of the data at 
increasing resolution: (\ref{eqn:y14}) represents the coarsest resolution (the sum of all data points), 
whereas (\ref{eqn:y1}) and (\ref{eqn:y3}) represent the finest resolution, with (\ref{eqn:y12}) in between. 
Further, this representation suggests a reparameterization, from $\bmu=(\mu_1,\mu_2,\mu_3,\mu_4)$
to $\mu_{1:4}$ plus the binomial parameters $\bp=(\mu_{1:2}/\mu_{1:4}, \mu_1/\mu_{1:2}, \mu_3/\mu_{3:4})$, where 
$p_1$ controls lower-resolution changes in the mean vector $\mu$ and $p_2,p_3$ control higher resolution changes.

@@

Note that, in this reparameterization, $p_j=0.5 \, \forall j$ corresponds to the case of a constant mean vector,
and values of $p_j$ far from 0.5 correspond to large changes in $\mu$ (at some scale).
Thus estimating a spatially-structured $\bmu$ can be achieved by shrinkage estimation of $\bp$, with shrinkage towards $p_j=0.5$. Both \cite{Nowak2000Statistical} and \cite{Timmermann1999Multiscale} use purpose-built Bayesian models to
achieve this shrinkage, by introducing a prior distribution on elements of $\bp$ that is
a mixture of a point mass at 0.5 (creating shrinkage toward 0.5) 
and a Beta distribution. Here we take a different approach, reparameterizing $\alpha_j = \log(p_j/(1-p_j))$,
and then using adaptive shrinkage to shrink $\alpha_j$ towards 0.

To apply adaptive shrinkage we need an estimate $\hat{\alpha}_{j}$ and corresponding standard error $\hat{s}_j$
for each $j$. This boils down to estimating a log-odds ratio, and its standard error, which is a well-studied problem
(e.g. \cite{Gart1967Bias}). The main challenge is in dealing satisfactorily with cases where the mle for $\alpha_j$ is infinite. Our choice of estimates, based on results from \cite{Gart1967Bias}, are described in Appendix \ref{}.

The output of adaptive shrinkage is a posterior distribution on each $\alpha_j$. The simplest approach
to obtain estimates of $\bmu$  would be to estimate
$\alpha_j$ by its posterior mean, and then reverse the reparameterization. 
The resulting estimate of $\mu_t$ would be the exponential of the posterior
mean for $\log(\mu_t)$ (because each $\log_(\mu_t)$ is a linear combination of $\alpha_j$). [IS THIS TRUE?]
An alternative is to estimate the posterior mean using the Delta method; see Appendix \ref{app:reconstruction}.

%(For example, to go to $T=8$, we have $Y_{1:8} \sim Poi(\mu_{1:8})$, $Y_{1:4} | Y_{1:8} \sim \Bin(Y_{1:8}, \mu_{1:4}/\mu_{1:8})$, and then (\ref{eqn:y12})-(\ref{eqn:y3}) and their analogues replacing indices 1-4 with 5-8.)

%Note that, just as for the original parameterization $Y_j \sim \Poi(\mu_j)$, the likelihood for the parameters in this reparameterization factorizes into a product of four terms, one for each parameter, corresponding to the four distributions (\ref{eqn:y14})-(\ref{eqn:y3}).

@@

\subsection{Simulations}
We now seek to validate the advantages of our method, named ``SMASH'', via simulation studies. To thoroughly investigate the performance of SMASH, we consider the Gaussian and Poisson cases separately.

For the Gaussian case, we focus our attention mostly on mean estimation. Different test functions, sample sizes, signal-to-noise ratios (SNR) and variance functions were considered, to ensure that SMASH behaves well in a variety of settings. However, only a small portion of the results will be shown here to highlight key findings from the rather extensive simulation study. Full results from the study can be found in the Supplementary Materials.

@@

To demonstrate the robustness of SMASH, we now consider only homoskedastic errors. Due to the large number of methods available in the literature for this setting, we chose only a subset. The methods we considered in our simulation are discussed in detail in \cite{Antoniadis2001Wavelet}, and in particular Translation Invariant (TI) thresholding by \cite{Coifman1995Translationinvariant} performed the best for most of the test signals, where performance is measured by mean squared error (MSE). We also considered the popular Empirical Bayes shrinkage procedure by \cite{Johnstone2005Empirical}, called Ebayesthresh. Figure \ref{fig:gaus_homo} compares the mean integrated squared errors (MISEs) of TI-thresholding and Ebayesthresh to SMASH with two different options for the ``Spikes'' mean function, with a signal to noise ratio of 3 and sample size 1024. The default option in SMASH estimates the mean without any homoskedastic assumptions on the variance function, while the second options estimates the mean given the true variance function. We can clearly see that the performance of SMASH is not impacted by the lack of homoskedastic assumptions imposed on the variance function: it outperforms two of the most accurate wavelet denoising algorithms for i.i.d. Gaussian noise even when the i.i.d. assumption holds, and performs nearly as well as when the variance function is known, which is the best-case scenario for any denoising technique. These results also extend to other mean functions, SNRs and sample sizes (see Supplementary Materials for more details).

@@

Having shown that SMASH performs well in the homoskedastic case without any explicit assumptions on the Gaussian noise, we now demonstrate its performance gain when the errors are heteroskedastic. For the sake of presentation, figure \ref{fig:gaus_hetero} highlights only the results from SMASH, TI-thresh (with various choices of variance estimation) and EbayesThresh for just two sets of test functions: the ``Spikes'' mean function with the ``Clipped Blocks'' variance function and the ``Corner'' mean function with ``Doppler'' variance function, both with SNRs of 3 and sample sizes of 1024. Nevertheless, similar results hold in general for different mean and variance functions, SNRs, and sample sizes (see Supplementary Materials for more details).

@@

It should also be noted that providing the true variance to SMASH does not result in substantial gains compared to the case when SMASH estimates its own variance function for most of the test functions (results shown in the supplementary materials). Hence, we can reasonably assume that SMASH does a good job of variance estimation, although this claim relies heavily on the variance function at hand. For example, the Bumps variance function (see appendix ?) is extremely difficult to estimate, and SMASH will perform much better when provided with the true variance function.

@@

    \caption{Mean (a) and variance (b) functions used in the simulation study comparing SMASH against MFVB.}

@@

In the first scenario, the number of data points is not a power of two, nor are the point equally spaced. To deal with these complications, we adapted and modified the standard symmetric extension procedure commonly used in wavelet settings. We first mirrored the data about the right edge and extract the first $2^{\lfloor\log_2(2n)\rfloor}$ sample points. This ensures that the number of data points in the new ``dataset'' is a power of two, and the mean curve would be continuous at the right edge. To further ensure that the input to the Gaussian denoising method is periodic, we then reflected the new dataset about the right edge and used this as the final input. To obtain the original mean and variance functions, we extracted the first $n$ points from the outputs (mean and variance) of our denoising technique. Since the data points are not evenly spaced, we took the simplest approach and applied our method treating the observations as if they were evenly spaced. This approach is not only intuitively appealing, but can also be considered a formal treatment of unequally spaced data in traditional wavelet settings (see eg. \cite{Sardy1999Wavelet}). Evaluation of MSE at the 201 equally spaced points is then based on simple linear interpolation between the estimated points. Tables \ref{table:mfvb_comp} display the MSEs over 100 independent runs for each scenario.

@@

\caption{MSEs of MFVB and SMASH for two simulation scenarios}

@@

Thus far, we have demonstrated that our method does a good job of mean and variance estimation for a variety of situations in the Gaussian case. As such, we now turn our attention to the Poisson case. For this simulation study, we considered different test functions and Poisson intensities for a given sample size of $n=1024$ as in \cite{Timmermann1999Multiscale}, \cite{Fryzlewicz2004HaarFisz} and \cite{Besbeas2004Comparative}, over 100 independent runs. Figure \ref{fig:pois_sim} compares the MISE of SMASH with Haar-Fisz (\cite{Fryzlewicz2004HaarFisz}) and BMSM (\cite{Kolaczyk1999Bayesian}), which are two popular and accurate Poisson denoising techniques, for the ``Bursts'' function with (min,max) intensities of (0.01,3) and (1/8,8). The low intensities are of primary interest in genomic applications, and so we focus our attention to these. Complete results, including a (min,max) intensity of (1/128, 128), are included in appendix \textcolor{red}{where?}. From figure \ref{fig:pois_sim}, it is clear that SMASH outperforms both methods, and the claim holds true in general. As an exception, Haar-Fisz outperforms both SMASH and BMSM for certain test functions with a (min,max) intensity of (1/128,128) when the asmyptotic variance of 1 is assumed in the Gaussian wavelet thresholding stage of the Haar-Fisz algorithm, but underperforms when the variance is estimated from the data instead. Besides the somewhat strong assumption of unit variance, the inconsistent performance of Haar-Fisz could also be attributed to the choice of the primary resolution level used, which can substantially affect the performance of Haar-Fisz in general. Here we analyzed primary resolution levels of 4, 5, 6 and 7, similar to \cite{Besbeas2004Comparative}. On the other hand, SMASH outperforms BMSM consistently, even though our method is based on the same likelihood factorization used in the latter. We can thus conclude that the choice of ASH as the shrinkage procedure (with the necessary likelihood approximations) is superior to that used in BMSM. More importantly, using this formulation in SMASH also makes it easily extensible to multiple samples in the context of a (generalized) linear model, something we will briefly describe in the Discussion section.
\begin{figure}
\centering
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{violin_pois_1.pdf}
        \caption{}
        \label{fig:pois_1}
    \end{subfigure}
		\hfill
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{violin_pois_8.pdf}
        \caption{}
        \label{fig:pois_8}
    \end{subfigure}
    \caption{Violin plots of MISEs for various methods. a) corresponds to a (min,max) intensity of (0.01,3), and b) corresponds to a (min,max) intensity of (1/8,8)}
    \label{fig:pois_sim}
\end{figure}

@@

>>>>

Methods for smoothing the variance of Gaussian data, or even for smoothing the mean in the presence of heteroskedastic errors, are uncommon, particularly compared with the well-studied homoskedastic case. Previous non-wavelet-based work includes \cite{Fan1998Efficient}, who estimated the variance by smoothing the squared residuals using local polynomial smoothing, \cite{Brown2007Variance}, who employed difference-based kernel estimators, and \cite{Menictas2015Variational}, who developed a Bayesian mean field variational methodology for both mean and variance estimation. The only wavelet-based method for variance estimation we have found is \cite{Cai2008Adaptive}, which applied a wavelet thresholding approach to first order differences. Unfortunately software implementation is only available in \cite{Menictas2015Variational}.

@@

%In this paper we use \ash~to tackle several signal denoising problems,  including Poisson smoothing,
%and smoothing both the mean and variance of heteroskedastic Gaussian data. 
Shrinkage methods are widely-used in signal denoising applications, because signal denoising can be accurately and conveniently achieved by shrinkage in a transformed (e.g.~wavelet) domain \cite{Donoho1994Ideal}. 
Commonly-used shrinkage methods include both
simple thresholding rules \cite{Coifman1995Translationinvariant, Donoho1994Ideal, donoho95} and EB methods \cite{Johnstone2005Empirical, Clyde2000Flexible}.
Being an EB method, \ash~has much in common with these previous EB methods, but generalizes them in two ways: first, 
\ash~allows more flexibility in the underlying distribution $g$ than the Laplace or spike-and-slab
distributions used in \cite{Johnstone2005Empirical, Clyde2000Flexible}; second, \ash~allows for variations in precision in the transformed observations (e.g. wavelet coefficients). The latter property is particularly important for
the Poisson and heteroskedastic Gaussian settings we consider here.

@@

The above outlines the basic strategy, but there are some important
 additional implementational details:
\begin{itemize}
\item Rather than use a single wavelet transform, we use the ``non-decimated" wavelet transform, which treats the observations as coming
from a circle, rather than a line, and averages results over all $T$ possible rotations of the data. Although not always necessary, this is a standard trick to reduce artifacts that can occur near discontinuities in the underlying signal (eg. \cite{Coifman1995Translationinvariant}). 
%COMMENT: WHat about reflection before doing this? ans: all the test fns are periodic
%DO OTHER EB APPROACHES CONSIDER THE NDWT? yes ebayesthresh software discusses this option
\item The non-decimated wavelet transform yields $T$ wavelet coefficients (transformed values of $\bY$) 
at each of $J=\log_2(T)$ resolution levels.
We apply ash separately to the wavelet coefficients at each resolution level, so that
a different distribution $g$ for the $(\tmu_j)$ is estimated for each resolution. This is the usual way that
EB approaches are applied in this context (e.g.~\cite{Johnstone2005Empirical}) and indeed is crucial because
the underlying distribution $g$  will vary with resolution (because smoothness of $\bmu$ will vary with resolution).  
\item Although we have presented the wavelet transform as a matrix multiplication, which is an $o(T^2)$ operation, in practice both the wavelet transform and the inverse transform are implemented using
efficient algorithms \cite{Beylkin1992Ontherepresentation,Coifman1995Translationinvariant}, implemented in the {\tt wavethresh} package \cite{Nason_wavethresh}, taking only $O(T\log T)$ operations. 
\end{itemize}
%\item Ash yields a posterior distribution for each $\tmu_t$. A simple approach would
%be to use the posterior mean for $\tmu_t$ and then invert the wavelet transform; however we instead use the delta method?

@@

Although we assumed variances to be known above, the problem of variance estimation is itself non-trivial. 
A common approach is to assume that the variance is constant ($\sigma_t=\sigma$), and that changes in the adjacent means $\mu_t-\mu_{t+1}$ are negligible, and to use an unbiased estimate for $\sigma$ such as
\begin{equation} \label{eq:constvar}
\hat{\sigma}^2 = (1/2T) \sum_{t=1}^{T} (Y_t-Y_{t-1})^2
\end{equation}
with $Y_0:=Y_T$.
Here we make the more flexible assumption that the variance function is spatially structured, and use wavelet shrinkage to estimate it. Our approach is similar in spirit to \cite{Cai2008Adaptive}, but we jointly estimate the mean and variance functions whereas they estimate only the variance function.

@@

To extend this to $T=4$, introduce the notation $v_{i:j}$ to denote, for any vector $v$, 
the sum $\sum_{t=i}^j v_t$.  Then 
\begin{align} 
Y_{1:4} & \sim \Poi(\mu_{1:4}) \label{eqn:y14} \\
Y_{1:2} | Y_{1:4} & \sim \Bin(Y_{1:4}, \mu_{1:2}/\mu_{1:4})  \label{eqn:y12} \\
Y_1 | Y_{1:2} & \sim \Bin(Y_{1:2}, \mu_1/\mu_{1:2}) \label{eqn:y1} \\ 
Y_3 | Y_{3:4} & \sim \Bin(Y_{3:4}, \mu_3/\mu_{3:4}). \label{eqn:y3}
\end{align} 
Together these models are exactly equivalent to $Y_j \sim \Poi(\mu_j)$, and
they decompose the overall distribution $Y_1,\dots,Y_4$ into parts involving aspects of the data at 
increasing resolution: (\ref{eqn:y14}) represents the coarsest resolution (the sum of all data points), 
whereas (\ref{eqn:y1}) and (\ref{eqn:y3}) represent the finest resolution, with (\ref{eqn:y12}) in between. 
Further, this representation suggests a reparameterization, from $\bmu=(\mu_1,\mu_2,\mu_3,\mu_4)$
to $\mu_{1:4}$ plus the binomial parameters $\bp=(\mu_{1:2}/\mu_{1:4}, \mu_1/\mu_{1:2}, \mu_3/\mu_{3:4})$, where 
$p_1$ controls lower-resolution changes in the mean vector $\mu$ and $p_2,p_3$ control higher resolution changes.

@@

Note that, in this reparameterization, $p_j=0.5 \, \forall j$ corresponds to the case of a constant mean vector,
and values of $p_j$ far from 0.5 correspond to large changes in $\mu$ (at some scale).
Thus estimating a spatially-structured $\bmu$ can be achieved by shrinkage estimation of $\bp$, with shrinkage towards $p_j=0.5$. Both \cite{kolaczyk1999wavelet} and \cite{Timmermann1999Multiscale} use purpose-built Bayesian models to
achieve this shrinkage, by introducing a prior distribution on elements of $\bp$ that is
a mixture of a point mass at 0.5 (creating shrinkage toward 0.5) 
and a Beta distribution. Here we take a different approach, reparameterizing $\alpha_j = \log(p_j/(1-p_j))$,
and then using adaptive shrinkage to shrink $\alpha_j$ towards 0.

To apply adaptive shrinkage we need an estimate $\hat{\alpha}_{j}$ and corresponding standard error $\hat{s}_j$
for each $j$. This boils down to estimating a log-odds ratio, and its standard error, which is a well-studied problem
(e.g. \cite{Gart1967Bias}). The main challenge is in dealing satisfactorily with cases where the maximum likelihood estimator (MLE) for $\alpha_j$ is infinite. Our choice of estimates, based on results from \cite{Gart1967Bias}, are described in Appendix \ref{}.

The output of adaptive shrinkage is a posterior distribution on each $\alpha_j$. The simplest approach
to obtain estimates of $\bmu$  would be to estimate
$\alpha_j$ by its posterior mean, and then reverse the reparameterization. 
The resulting estimate of $\mu_t$ would be the exponential of the posterior
mean for $\log(\mu_t)$ (because each $\log(\mu_t)$ is a linear combination of $\alpha_j$).
Alternatively we can estimate $\mu_t$ by approximating its posterior mean using the Delta method; see Appendix \ref{app:reconstruction}.
Both methods are implemented in our software; for the results here we use the latter method 
to be more comparable with previous approaches that estimate $\mu$ on the raw scale rather than $\log$ scale.
%Note that, just as for the original parameterization $Y_j \sim \Poi(\mu_j)$, the likelihood for the parameters in this reparameterization factorizes into a product of four terms, one for each parameter, corresponding to the four distributions (\ref{eqn:y14})-(\ref{eqn:y3}).

@@

@@

\subsection{Simulations}

We conducted extensive simulation studies to compare the performance of our method with existing approaches.

For the Gaussian case, we focus on mean estimation, initially with homoskedastic errors. We modelled our simulation study after
\cite{Antoniadis2001Wavelet}, using many of the same test functions, a variety of sample sizes, two different signal to noise ratios (SNRs), and including
the best-performing methods from their comparison. In particular Translation Invariant (TI) thresholding (\cite{Coifman1995Translationinvariant}) performed the best for most of the test signals, where performance is measured by mean squared error (MSE). We also considered the popular Empirical Bayes shrinkage procedure (\cite{Johnstone2005Empirical}), named Ebayesthresh. Figure \ref{fig:gaus_homo} compares the mean integrated squared errors (MISEs) of TI-thresholding and Ebayesthresh to SMASH with two different options for the ``Spikes'' mean function, with a signal to noise ratio of 3 and sample size 1024. The default version of SMASH estimates the mean without any homoskedastic assumptions on the variance function, while the other version of SMASH estimates the mean given the true variance function. We can clearly see that the performance of SMASH is not impacted by the lack of homoskedastic assumptions imposed on the variance function: it outperforms two of the most accurate wavelet denoising algorithms for i.i.d. Gaussian noise even when the i.i.d. assumption holds, and performs nearly as well as when the variance function is known, which is the best-case scenario for any denoising technique. These results also extend to other mean functions, SNRs and sample sizes (see Supplementary Materials for more details).
 Different test functions, sample sizes, signal-to-noise ratios (SNR) and variance functions were considered, to ensure that SMASH behaves well in a variety of settings. However, only a small portion of the results will be shown here to highlight key findings from the extensive simulation study. Full results from the study can be found in the Supplementary Materials.

We focus initially on estimation of a Gaussian mean with homoskedastic errors. Due to the large number of methods available in the literature for this setting, we chose only a subset.

@@

Having shown that SMASH performs well in the homoskedastic case without any explicit assumptions on the Gaussian noise, we now demonstrate its performance gain when the errors are heteroskedastic. For the sake of presentation, Figure \ref{fig:gaus_hetero} highlights only the results from SMASH, TI-thresh (with various choices of variance estimation) and EbayesThresh for just two sets of test functions: the ``Spikes'' mean function with the ``Clipped Blocks'' variance function and the ``Corner'' mean function with ``Doppler'' variance function, both with SNRs of 3 and sample sizes of 1024. Nevertheless, similar results hold in general for different mean and variance functions, SNRs, and sample sizes (see Supplementary Materials for more details).

@@

It should also be noted that providing the true variance to SMASH does not result in substantial gains compared to the case when SMASH estimates its own variance function for most of the test functions (results shown in the Supplementary Materials). Hence, we can reasonably assume that SMASH does a good job of variance estimation, although this claim relies heavily on the variance function at hand. For example, the Bumps variance function (see Appendix ??) is extremely difficult to estimate, and SMASH will perform much better when provided with the true variance function.

@@

    \caption{Mean (a) and variance (b) functions used in the simulation study comparing SMASH against MFVB. These functions correspond to mean and standard deviation functions (A) from \cite{Menictas2015Variational}.}

@@

In the first scenario, the number of data points is not a power of two, nor are the point equally spaced. To deal with these complications, we adapted and modified the standard symmetric extension procedure commonly used in wavelet settings. We first mirrored the data about the right edge and extract the first $2^{\lfloor\log_2(2n)\rfloor}$ sample points. This ensures that the number of data points in the new ``dataset'' is a power of two, and the mean curve would be continuous at the right edge. To further ensure that the input to the Gaussian denoising method is periodic, we then reflected the new dataset about the right edge and used this as the final input. To obtain the original mean and variance functions, we extracted the first $n$ points from the outputs (mean and variance) of our denoising technique. Since the data points are not evenly spaced, we took the simplest approach and applied our method treating the observations as if they were evenly spaced. This approach is not only intuitively appealing, but can also be considered a formal treatment of unequally spaced data in traditional wavelet settings (eg. \cite{Sardy1999Wavelet}). Evaluation of MSE at the 201 equally spaced points is then based on simple linear interpolation between the estimated points. Table \ref{table:mfvb_comp} displays the MSEs over 100 independent runs for each scenario.

@@

\caption{Compares the performance (in MSE) of SMASH and MFVB for two simulation scenarios, with the mean and variance functions as shown in \ref{fig:mfvb_fn}. SMASH is comparable to MFVB in mean estimation and outperforms the latter in variance estimation in Scenario 1, where the data are not equally spaced and not a power of 2. SMASH outperforms MFVB in both mean and variance estimation in Scenario 2, where the data are equally spaced and a power of 2.}

@@

Thus far, we have demonstrated that our method does a good job of mean and variance estimation for a variety of situations in the Gaussian case. As such, we now turn our attention to the Poisson case. For this simulation study, we considered different test functions and Poisson intensities for a given sample size of $n=1024$ as in \cite{Timmermann1999Multiscale}, \cite{Fryzlewicz2004HaarFisz} and \cite{Besbeas2004Comparative}, over 100 independent runs. Figure \ref{fig:pois_sim} compares the MISE of SMASH with Haar-Fisz (\cite{Fryzlewicz2004HaarFisz}) and BMSM (\cite{Kolaczyk1999Bayesian}), which are two popular and accurate Poisson denoising techniques, for the ``Bursts'' function with (min,max) intensities of (0.01,3) and (1/8,8). The low intensities are of primary interest in genomic applications, and so we focus our attention to these. Complete results, including a (min,max) intensity of (1/128, 128), are included in Supplementary Materials. It is clear from Figure \ref{fig:pois_sim} that SMASH outperforms both methods, and the claim holds true in general. As an exception, Haar-Fisz outperforms both SMASH and BMSM for certain test functions with a (min,max) intensity of (1/128,128) when the asmyptotic variance of 1 is assumed in the Gaussian wavelet thresholding stage of the Haar-Fisz algorithm, but underperforms when the variance is estimated from the data instead. Besides the somewhat strong assumption of unit variance, the inconsistent performance of Haar-Fisz could also be attributed to the choice of the primary resolution level used, which can substantially affect the performance of Haar-Fisz in general. Here we analyzed primary resolution levels of 4, 5, 6 and 7, similar to \cite{Besbeas2004Comparative}. On the other hand, SMASH outperforms BMSM consistently, even though our method is based on the same likelihood factorization used in the latter. We can thus conclude that the choice of ASH as the shrinkage procedure is superior to that used in BMSM. More importantly, using this formulation in SMASH also makes it easily extensible to multiple samples in the context of a (generalized) linear model, something we will briefly describe in the Discussion section.
\begin{figure}
\centering
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{violin_pois_1.pdf}
        \caption{}
        \label{fig:pois_1}
    \end{subfigure}
		\hfill
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{violin_pois_8.pdf}
        \caption{}
        \label{fig:pois_8}
    \end{subfigure}
    \caption{Compares various approaches for denoising Poisson data for the ``Bursts'' test function where a) corresponds to a (min,max) intensity of (0.01,3), and b) corresponds to a (min,max) intensity of (1/8,8). Violin plots of MISEs for (from bottom to top) SMASH, BMSM and Haar-Fisz are shown in the figure. Smaller MISE implies better performance; dashed green line indicates the median MISE for SMASH. We note the performance gain of SMASH over both BMSM and Haar-Fisz.}
    \label{fig:pois_sim}
\end{figure}

@@

    \caption{Results from fitting SMASH to the motorcycle acceleration data discussed in \cite{silverman1985some}. Both the mean curve (solid black line) and the variance around the mean curve (dashed red line) are shown. We note that SMASH does a good job of capturing both the mean as well as the variation in the data without requiring and parameter tuning.}
